### 2021-01-09 paper/paper.tex

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in
\usepackage{ijcai21}
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\urlstyle{same}

% \usepackage{showframe}

\newtheorem{proposition}{Proposition}

\usepackage{complexity}

\pdfinfo{
/TemplateVersion (IJCAI.2021.0)
}

\title{Solving Graph Homomorphism and Subgraph Isomorphism Problems Faster Through Neighbourhood
Maximum Clique Constraints}

\author{
Ciaran McCreesh$^1$
\and
Sonja Kraiczy$^2$
\affiliations
$^1$University of Glasgow\\
$^2$University of Oxford\\
\emails
ciaran.mccreesh@glasgow.ac.uk
}

\newcommand{\neighbourhood}{\operatorname{N}}
\newcommand{\vertexset}{\operatorname{V}}
\newcommand{\degree}{\operatorname{deg}}
\newcommand{\nds}{\operatorname{S}}

\begin{document}

\maketitle

\begin{abstract}
    We can find graph homomorphisms faster through clique constraints.
\end{abstract}

\section{Introduction}

A \emph{graph homomorphism} is a function between two graphs that maps adjacent vertices to adjacent
vertices. An injective graph homomorphism is more commonly known as a \emph{subgraph isomorphism},
whilst a \emph{locally injective homomorphism} is one which is injective when restricted to any
individual vertex and its neighbourhood \cite{DBLP:journals/csr/FialaK08}. Finding any of these three kinds of homomorphism between
two given graphs is \NP-complete.  The subgraph isomorphism problem arises in many practical
applications, including ??, and as such there has been a lot of research into finding algorithms
which perform well empirically. Most approaches are based either upon very fast but simple
backtracking algorithms,
\cite{DBLP:journals/pami/CordellaFSV04,DBLP:journals/bmcbi/BonniciGPSF13,DBLP:conf/gbrpr/CarlettiFSV17}
which often but not always perform well on very easy instances, or upon constraint programming
algorithms
\cite{DBLP:journals/constraints/ZampelliDS10,DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/ArchibaldDHMP019},
which have higher startup costs but that perform vastly better on harder instances and much more
consistently on easy instances \cite{DBLP:journals/jair/McCreeshPST18,DBLP:conf/gbrpr/Solnon19}. The
current state of the art is the Glasgow Subgraph Solver \cite{DBLP:conf/gg/McCreeshP020}, which is a
dedicated constraint programming solver for subgraph-finding problems. Much of its performance comes
from inference strategies based upon degrees and neighbourhood degree sequences
\cite{DBLP:journals/constraints/ZampelliDS10}, counting paths between vertices
\cite{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, and cardinality reasoning
\cite{DBLP:journals/ai/Solnon10}, which can be used to eliminate many infeasible candidate
assignments without search.

One might hope that this inference would also be helpful for other graph homomorphism problem
variants. Indeed, this paper proves that many, but not all, of these strategies are also valid in
the locally injective case, and that a weakened form of one other strategy is valid for all
homomorphisms. However, we also prove that \emph{none} of the other strategies are valid for finding
homomorphisms where there is no injectivity requirement. Finally, we introduce a new filtering
technique that is based upon finding the maximum clique in the neighbourhood of each vertex, and
prove that this technique is valid even in the general case. Although this new filtering technique
involves solving a large number of \NP-complete problems, we demonstrate that it is effective in
practice, particularly for the non-injective problem where filtering can solve instances ??many
hundreds of times faster in aggregate. This shows, for the first time, that modern constraint
programming techniques can be practical for less constrained graph homomorphism finding problems;
previous algorithmic approaches have focused instead upon worst-case computational bounds
\cite{DBLP:journals/csr/FialaK08,DBLP:journals/ipl/Rzazewski14,DBLP:journals/tcs/ChaplickFHPT15},
whose practical utility has yet to be demonstrated. The empirical effectiveness of our results is
especially important because many current applications use subgraph isomorphism solvers only because
they perform well using off the shelf solvers, rather than because they exactly match ideal domain
requirements ??cite.

\section{Background and Theory}

Let $G$ and $H$ be graphs. Let $v \in \vertexset(G)$ be a vertex of $G$. The \emph{(open)
neighbourhood} of $v$, written $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$ not
including $v$ itself, whilst the closed neighbourhood of $v$, written $\neighbourhood_G[v]$, is the
neighbourhood plus $v$. The \emph{degree} of a vertex, $\degree_G(v)$, is the cardinality of its open
neighbourhood. Given a vertex set $S \subseteq \vertexset(G)$, the subgraph \emph{induced by} $S$,
written $G[S]$, is the subgraph of $G$ with only the vertices in $S$ together with all the edges
between them. A \emph{clique} is a subgraph where every vertex is adjacent to every other in the
subgraph.

A \emph{homomorphism} from $G$ to $H$ is a function mapping vertices of $G$ to vertices of $H$, such
that adjacent vertices in $G$ are mapped to adjacent vertices in $H$. A homomorphism $h$ is
\emph{locally injective} if for every vertex $w$, the restriction of $h$ to $G[N_G[w]]$ is
injective; if $h$ is injective globally we call it a \emph{(non-induced) subgraph isomorphism}.

?? loops, be really careful with this

?? CSP, hom as CSP

An \emph{implied constraint} is a constraint that can be added to a CSP without altering its set of
solutions. ?? Is it best wording all of this in terms of implied constraints?

\begin{proposition}For the problem of finding a subgraph isomorphism $i$ from a graph $G$ to a graph
    $H$, the following constraints are implied for any vertex $v \in \vertexset(G)$:
    \begin{enumerate}
        \item We cannot map $v$ to a vertex of lower degree, $\degree(v) \le \degree(i(v))$.
        \item Furthermore, ??NDS.
    \end{enumerate}
\end{proposition}

?? Show that these work for locally injective but not homomorphism

\begin{proposition}For the problem of finding a subgraph isomorphism $i$ from a graph $G$ to a graph
    $H$, the following constraints are implied for any pair of vertices $v, w \in \vertexset(G)$:
    \begin{enumerate}
        \item The distance between $v$ and $w$ is at most the distance between $i(v)$ and $i(w)$.
        \item For any pair of vertices $v, w \in \vertexset(G)$, if there are at least $k$ simple
            paths of length exactly $\ell$ between $v$ and $w$, then there must be at least $k$
            simple paths of length exactly $\ell$ between $i(v)$ and $i(w)$.
    \end{enumerate}
\end{proposition}

?? This fourth one is hard. The third one actually works for homomorphisms.

?? Corollary: $G^d$ graphs.

?? Clique, be very careful with target loops. What we want to prove is, if the maximum clique in the
neighbourhood of a pattern vertex $p$ has $n$ vertices, and if the target graph has no loops, then
$p$ cannot be mapped to any target vertex $t$ unless $t$ is contained in a clique of at least $n$
vertices.

?? The point is we can solve clique decision much faster than general homomorphism, and we only need
to do it once per target vertex.

\section{Design and Implementation}

We use the subgraph isomorphism code in the Glasgow Subgraph Solver \cite{DBLP:conf/gg/McCreeshP020}
as a starting point.  We implemented distance filtering for the basic homomorphism problem, using
the existing supplemental graphs framework that the solver uses for subgraph isomorphism to create
pseudo-adjacency constraints for vertices that are distance two away from each other; this was
routine.  We also implemented clique constraints, initially using the Glasgow Subgraph Solver's
existing maximum clique solver to perform domain filtering. Preliminary experiments showed that a
na{\"\i}ve approach, which begins by calculating the maximum clique size for the neighbourhood of
each pattern vertex and each target vertex, would add as much as three minutes of preprocessing time
to some problem instances which could otherwise be solved in a few seconds. We therefore use a more
intelligent approach, which we now describe.

\subsection{Optimising Maximum Clique Solver Calls}

We calculate the maximum clique size for the neighbourhood of a first pattern vertex $p_1$.  Having
found a maximum clique $\omega_1$ for a vertex $p$, for every other vertex $q \in \omega_1$, we
remember that the maximum clique for $q$'s neighbourhood has at least $|\omega_1|$ vertices. We then
proceed to find a maximum clique $\omega_2$ for a second pattern vertex $p_2$; however, if $p_2$ was
a member of $\omega_1$, we initialise the clique solver with a lower bound of $n$ rather than
starting from scratch. We then find the maximum clique size for a third pattern vertex $p_3$, using
whichever lower bound of $|\omega_1|$ and $|\omega_2|$ is better (if both are valid), and so on.

We then move on to the target vertices, using a similar caching routine. Rather than calculating a
clique size for every single target vertex, we only calculate a value for target vertices which are
present in at least one variable's domain, after other unary constraints have been applied.
(Depending upon the problem variant, implied constraints like degree and neighbourhood degree
sequence can eliminate some target vertices from the domains of all pattern variables.)
Additionally, we do not require the maximum clique solver to run to completion and guarantee that it
has found a maximum clique. Instead, we allow it to stop as soon as it has found a clique with as
many vertices as the largest pattern clique. This is useful in practice because it may be very hard
to decide whether a particular target vertex has neighbourhood clique size of, say, 15 or 16, but if
the largest pattern vertex has a neighbourhood clique size of only 5 then this is irrelevant.

?? Why this works

\subsection{Proof Logging}

?? How to log proofs

\section{Experiments}

\begin{figure*}[p]
    \,\hfill\begin{tabular}{c@{}c@{}c@{}}
    \includegraphics{gen-graph-cumulatives.pdf}
        &
    \includegraphics{gen-graph-cumulatives-differences.pdf}
        &
    \includegraphics{gen-graph-cumulatives-aggregate.pdf}
        \\[5mm]
    \includegraphics{gen-graph-cumulatives-hom.pdf}
        &
    \includegraphics{gen-graph-cumulatives-local.pdf}
        &
    \includegraphics{gen-graph-cumulatives-si.pdf}
    \end{tabular}\hfill\,

    \caption{On the top left, the cumulative number of instances solved over time for the three
    problem variants, with and without clique filtering, and with and without distance filtering for
    the homomorphism problem. The remaining plots re-display this data, as follows. The three plots
    on the bottom row zoom in on the cumulative number of instances solved, for the homomorphism
    problem on the left, the locally injective homomorphism problem in the centre, and the subgraph
    isomorphism problem on the right. The top centre plot shows the additional number of instances
    solved at any given time when using the new forms of filtering for all problem variants, and the
    top right plot shows the aggregate speedups from each form of filtering.}
\end{figure*}
\begin{figure*}[p]
    \,\hfill\begin{tabular}{c@{\hspace{4mm}}c@{\hspace{4mm}}c@{\hspace{4mm}}c@{\hspace{4mm}}}
        \includegraphics{gen-graph-scatter-hom-c.pdf}
        &
        \includegraphics{gen-graph-scatter-hom-cd.pdf}
        &
        \includegraphics{gen-graph-scatter-local.pdf}
        &
        \includegraphics{gen-graph-scatter-si.pdf}
    \end{tabular}\hfill\,

    \caption{Looking at the effects of additional filtering on an instance by instance basis, for
    homomorphism with just clique filtering and with both clique and distance filtering, and for the
    other two variants with clique filtering. Each point represents one instance, the vertical axis
    is the runtime with filtering in ms, and the horizontal axis is the runtime without filtering in
    ms (and so points below the diagonal are speedups). Points on the outer axes are timeouts. The
    different point styles show the different families of instance from the benchmark set, and
    illustrate that in each the filtering is broadly useful rather than being specific to a single
    kind of application.}
\end{figure*}

We now evaluate this approach empirically.

?? How good is it

?? Time on cliques

?? Whether cliques are hard

\section{Conclusion and Future Work}

?? Neighbourhood colourability? Cores?

\section*{Acknowledgements}

This work was supported by the Engineering and Physical Sciences Research Council [grant number
EP/P026842/1].

\bibliographystyle{named}
\bibliography{paper}

\end{document}

% vim: set tw=100 spell spelllang=en : %